Dominance Regions for Rank Two Cluster Algebras

### 2022-05-01 dominance_regions.tex
Minor intro edits

  Our goal in this note is to explicitly compute the possible bases which are pointed with respect to the embedding $\cA(b,c)\into\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ for every $k\in\ZZ$.
  That is, a basis element with \emph{$\bfg$-vector} $\lambda=(\lambda_0,\lambda_1)\in\ZZ^2$ when expanded in terms of $\{x_0,x_1\}$ can be written in the form
  \begin{equation}
    \label{eq:pointed}
    x_0^{\lambda_0}x_1^{\lambda_1}\sum\limits_{\alpha_0,\alpha_1 \ge 0} \rho_{\alpha_0,\alpha_1} x_0^{-b\alpha_0} x_1^{c\alpha_1}
  \end{equation}
  with $c_{0,0}=1$.
  Being \emph{pointed} means an analogous structure reproduces after expanding in terms of any pair $\{x_k,x_{k+1}\}$.
  %The collection of all $\bfg$-vectors $\lambda\in\ZZ^2$ parametrize the elements of any pointed basis.

  Important examples of pointed elements are the \emph{cluster monomials} of $\cA(b,c)$, i.e. the elements that can be written in the form $x_k^{\alpha_k}x_{k+1}^{\alpha_{k+1}}$ for some $k\in\ZZ$ and $a_k,a_{k+1}$ non-negative integers.
  Indeed, by \cite[Lemma 3.4.12]{qin} cluster monomials are part of any pointed basis in any cluster algebra.
  We achieve the same conclusion in our setting by elementary calculations (c.f. Lemma~\ref{le:cluster monomials}).

@@

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance region $\cP_\lambda$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the polygon consisting of those $\mu\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

  A number of bases for rank two cluster algebras with various defining properties have been constructed.
  A natural starting basis is the so-called \emph{greedy basis} in which all $c_{\alpha_0,\alpha_1}$ are non-negative and these coefficients are chosen to be minimal.
  It is known that this choice coincides with the theta basis \cite{GHKK, CGMMRSW}.
  The \emph{standard monomial bases} of $\cA(b,c)$ come from natural identifications with the subalgebras of $\ZZ[x_0^{\pm1},x_1^{\pm1}]$ generated by $x_{k-1}$, $x_k$, $x_{k+1}$, and $x_{k+2}$ for $k\in\ZZ$.
  However, these bases depend on $k$ and the triangular basis is defined by a triangular transition with the stipulation that the basis does not depend on the choice of starting cluster.
  It would be useful to explicitly (or combinatorially) compute the transitions between these bases.
  However, explicit calculations reveal that the $\bfg$-vectors actually appearing in these transitions are far smaller than the entire dominance polygon.

@@

>>>>

  Our goal in this note is to explicitly compute the possible bases which are pointed with respect to the embedding $\cA(b,c)\into\ZZ[x_k^{\pm1},x_{k+1}^{\pm1}]$ for every $k\in\ZZ$.
  That is, a basis element with \emph{$\bfg$-vector} $\lambda=(\lambda_0,\lambda_1)\in\ZZ^2$ when expanded in terms of $\{x_0,x_1\}$ can be written in the form
  \begin{equation}
    \label{eq:pointed}
    x_0^{\lambda_0}x_1^{\lambda_1}\sum\limits_{\alpha_0,\alpha_1 \ge 0} \rho_{\alpha_0,\alpha_1} x_0^{-b\alpha_0} x_1^{c\alpha_1}
  \end{equation}
  with $\rho_{0,0}=1$.
  Being \emph{pointed} means an analogous structure reproduces after expanding in terms of any pair $\{x_k,x_{k+1}\}$.
  %The collection of all $\bfg$-vectors $\lambda\in\ZZ^2$ parametrize the elements of any pointed basis.

  Important examples of pointed elements are the \emph{cluster monomials} of $\cA(b,c)$, i.e. the elements that can be written in the form $x_k^{\alpha_k}x_{k+1}^{\alpha_{k+1}}$ for some $k\in\ZZ$ and $\alpha_k,\alpha_{k+1}$ non-negative integers.
  Indeed, by \cite[Lemma 3.4.12]{qin} cluster monomials are part of any pointed basis in any cluster algebra.
  We achieve the same conclusion in our setting by elementary calculations (c.f. Lemma~\ref{le:cluster monomials}).

@@

  When $bc\le3$, the cluster algebra $\cA(b,c)$ will be of finite-type and cluster monomials form its only pointed basis.
  We will therefore assume that $bc\ge4$.
  In this case we give an explicit description of the dominance relation among $\bfg$-vectors.
  Specifically, we show that the $\bfg$-vector $\lambda$ dominates the collection of $\bfg$-vectors of the form $\lambda+(b \alpha_0 ,c \alpha_1)$, $\alpha_0,\alpha_1\in\ZZ$, inside its \emph{dominance region} $\cP_\lambda$.
  %In this case, we compute the possible transitions between any two pointed bases.
  %The collection of $\bfg$-vectors which can appear in the deformation of a pointed basis is the \emph{dominance region}.
  \begin{theorem}
    \label{th:dominance inequalities}
    The dominance region $\cP_\lambda$ pointed at $\lambda=(\lambda_0,\lambda_1)$ is the polygon consisting of those $\mu=(\mu_0,\mu_1)\in\RR^2$ satisfying $\mu_0 \leq \lambda_0, \mu_1 \geq\lambda_1$, and the following inequalities:
    {
      \everymath={\displaystyle}
      \def\arraystretch{2.8}
      \[
        \begin{array}{rcccl}
          0 & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)+(\mu_1-\lambda_1) & \leq & \max\left(-c\lambda_0-\frac{b c+\sqrt{b c (b c-4)}}{2b}b\lambda_1,0\right)
          \\
          \min\left(-c\lambda_0,0\right) & \leq & \frac{b c-\sqrt{b c (b c-4)}}{2 b}(\mu_0-\lambda_0)-(\mu_1-\lambda_1) & \leq & 0
          \\
          0 & \leq &  -(\mu_0-\lambda_0)-\frac{b c-\sqrt{b c (b c-4)}}{2 c}(\mu_1-\lambda_1) & \leq & \max\left(\frac{b c+\sqrt{b c (b c-4)}}{2c}c\lambda_0+b\lambda_1,0\right)
          \\
          \min\left(b \lambda_1,0\right) & \leq & (\mu_0-\lambda_0) - \frac{b c-\sqrt{b c (b c-4)}}{2 c} (\mu_1-\lambda_1) & \leq & 0
        \end{array}
      \]
    }
  \end{theorem}

@@

  A number of bases for rank two cluster algebras with various defining properties have been constructed.
  A natural starting basis is the so-called \emph{greedy basis} in which all $\rho_{\alpha_0,\alpha_1}$ are non-negative and these coefficients are chosen to be minimal.
  It is known that this choice coincides with the theta basis \cite{GHKK, CGMMRSW}.
  The \emph{standard monomial bases} of $\cA(b,c)$ come from natural identifications of $\cA(b,c)$ with the subalgebras of $\ZZ[x_0^{\pm1},x_1^{\pm1}]$ generated by $x_{k-1}$, $x_k$, $x_{k+1}$, and $x_{k+2}$ for $k\in\ZZ$.
  However, these bases depend on $k$ and the triangular basis is defined by a triangular transition with the stipulation that the basis does not depend on the choice of starting cluster.
  It would be useful to explicitly (or combinatorially) compute the transitions between these and other known bases.
  However, explicit calculations reveal that the $\bfg$-vectors actually appearing in these transitions are far smaller than the entire dominance polygon.
